### 2009-12-16 demo/demo.tex
added slides.

\documentclass[xcolor=dvipsnames,professionalfonts,smaller,presentation]{beamer} 
\usecolortheme[named=Brown]{structure} 
\useoutertheme[subsection=false]{miniframes}
\usetheme{Malmoe}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[portuguese]{babel}
\usepackage[latin1]{inputenc}
\usepackage{alltt}
\usepackage{moreverb}
\usepackage{hyperref}
\usepackage{indentfirst}
\usepackage{lastpage}
\usepackage{url}
\usepackage{float}

\urldef{\mails}\path|ei05011@fe.up.pt,ei05028@fe.up.pt|
\title[A comparison of classification algorithms applied to the Adult Database \insertframenumber/\inserttotalframenumber]{A comparison of classification algorithms applied to the Adult Database}
\author{Flávio Cruz and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
\mails}
\date{\today}

\begin{document}

\frame{\titlepage}

\section[Outline]{}
\frame{\tableofcontents}

\section{Problem proposed by the Dataset}
\frame
{
}

\section{Dataset Characterization}
\frame
{
}

\section{Data Mining Problem}
\frame
{
}

\section{Algorithms Used}
\frame
{
}

\section{Conducted Experiments}
\frame
{
}

\section{Results}
\frame
{
}

\section{Conclusions}
\frame
{
}

\end{document}